---
title: "Movies"
output:
  html_document:
    theme: flatly
    toc: true
    toc_float: true
    highlight: zenburn
---

```{r echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(ggplot2)
library(pander)
library(scales)

movies <- read_csv("~/Fall 2024/Professional Projects/Data/movies.csv")
```


Movies have almost always been a go to topic to get conversations started, especially when meeting someone for the first time. It helps get a grasp on what others might be interested in, as well as possible movie recommendations. As I collected data for this, I tried to stick to more well known movies, but included movies from varying genres and ratings (G, PG, PG-13, R). Included in the data are 40 movies from each chosen genre, what each movie is rated, the audience score the movie has, and the success it had at the box office with it's worldwide gross.

## {.tabset .tabset-pills .tabset-fade}

### Genre
#### **Genre VS Box Office**



### Audience Score
#### **Audience Scores VS Box Office**



### Rating
#### **Rating VS Box Office**

```{r echo=FALSE, message=FALSE, warning=FALSE}
movies <- movies %>%
  mutate(gross = as.numeric(gsub("[^0-9]", "", gross)))

RatingMean <- movies %>% 
  group_by(rating) %>% 
  summarise(avg = mean(gross, na.rm = TRUE)) 

ggplot(RatingMean, aes(x = rating, y = avg, fill = rating))+
  geom_bar(stat = "identity", color = "black")+
    geom_text(aes(label = scales::dollar(avg)), vjust = -0.5, size = 3) + 
  scale_y_continuous(labels = scales::dollar_format()) +
  labs(
    title = "Average Gross by Movie Rating",
    x = "Movie Rating",
    y = "Average Gross (USD)",
    fill = "Rating") +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title = element_text(hjust = 0.5, size = 16, face = "bold"))

```

```{r}

```

```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(movies, aes(x = factor(rating), y = gross))+
  geom_dotplot(binaxis = "y", stackdir = "up", position = "dodge", dotsize = 0.5)+
   ylim(0, 500000000)+
  labs(
    title = "Worldwide Gross Success by Rating",
    x = "Movie Rating",
    y = "Worldwide Gross at the Box Office (USD)")+
    theme_minimal()

```




##