---
title: "Weather Predictor"
output:
  html_document:
    theme: flatly
    toc: true
    toc_float: true
    highlight: zenburn
---

```{r echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(ggplot2)
library(pander)
library(scales)
library(DT) 
library(mosaic)

IFweather <- read_csv("~/Fall 2024/Professional Projects/Data/weather.csv")
IFweather <- IFweather %>% 
  select(where(~ !all(is.na(.))))
```

## {.tabset .tabset-pills .tabset-fade}
### Idaho Falls Weather
Idaho Falls, Idaho, a place where, compared to many other places, there's ample snow, strong winds, and cold temperatures. The weather may not seem to always be predictable, but with the data that I was able to collect using [Weather Underground](https://www.wunderground.com/history/daily/us/id/idaho-falls/KIDIDAHO6/date/2024-1-6), I have put together an analysis comparing past and present weather patterns to help create a weather model predictor. 

### Data
```{r echo=FALSE, message=FALSE, warning=FALSE}
datatable(IFweather, options=list(lengthMenu = c(3,10,30)), extensions="Responsive")
```
##
